Development of Physics or HPC Optimisation of a parallel 2D lattice Boltzmann solver using GPUs/CUDA

### 2017-07-24 Thesis/Draft.tex
post present with g

\usepackage[utf8]{inputenc}
% Use the custom "cranfieldthesis" LaTeX style file. 
\usepackage{cranfieldthesis}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}

% By default, LaTeX uses a serif font - these are traditionally thought to be
% easier to read.   If you'd prefer sans-serif, please uncomment the 
% following line.
%\renewcommand{\familydefault}{\sfdefault}

@@

\section{Lattice Boltzmann Method}
The Lattice Boltzmann method (LBM) is a mesoscopic scale approach to CFD and was first introduced as a Lattice-Gas Automata for the Navier-Stokes Equation \cite{gasautomata}. It is used to describe a fluid based on probabilities using the Maxwell-Boltzmann equation in the fluid’s equilibrium state \cite{lbm_springer}. 
\par In this method, we do not consider the individual characteristics of each particle. By grouping particles together in a D2Q9 (nodes containing 9 particles for 2D problems - Fig \ref{fig:d2q9})  or in a D3Q19 (nodes containing 19 particles for 3D problems - Fig \ref{fig:d3q19}) model, we can analyse the behaviour of the particles collectively  \cite{lbm_springer}.

@@

\subsection{Multiphase flow}
Multiphase flows represent the simultaneous flow of materials in different states (phases). As such, multiphase flows have an enormous spectrum of representation and stand for interactions between gas/solid flows, liquid/solid flows or even liquid/liquid flows with different chemical properties \cite{multiphase}.\par
These simulations present challenging problems because of inherent difficulties during modelling and the importance of engineering applications. However, the Lattice Boltzmann method provides an alternative for these simulations due to it's relative simplicity when compared with traditional Navier-Stokes equation \cite{chenDoolenMP}.
\par These flows have several applications in the scientific and industrial community, ranging from pourous media fluid interactions \cite{pourous} to flows containing gas bubbles dispersed in liquids \cite{gasBubble}. Furthermore, almost every processing technology must take multiphase flows into account, including cavitating pumps, papermaking and many others \cite{multiphase}.
\par Several implementation strategies exist for Multiphase flows, however, for the purpose of this thesis, we will first follow the simple ''two-color'' approach first proposed by Gunstensen et al \cite{gunstensen}.

@@

\section{Previous parallelisation works}
This thesis is a continuation of work that started some years ago in Cranfield University. As such, the work from the previous students needs to be analysed. \par
In 2014, Tamás Józsa and Máté Szőke, adapted two different in-house C and C++ codes into one single C code unifying the advantages of each one of the two original codes \cite{jozsa_thesis,szoke_thesis}. Józsa then parallelised the critical parts of the C code using CUDA and ran tests on the Fermi GPU Cluster from Cranfield, achieving a three times speed-up in general, with a peak of 15 times speed-up \cite{jozsa_thesis}. \par
Szőke proposes a CPU parallelisation approach using Unified Parallel C, which is a Partitioned Global Address Space language\cite{szoke_thesis}. This means that it is possible to use shared memory to compute the solution. However, the author verified that a local memory-based approach (like MPI) provided the best results. He also compared the results obtained with the ones obtained by Józsa on the CUDA approach. They found that to achieve the same speed-up as that of a single GPU card, one needs an entire workstation (16 threads in the case of Astral) \cite{szoke_thesis}.\par
In 2015, Ádám Koleszár continued the work and further optimised the parallel version of the LBM method using CUDA \cite{koleszar_thesis}. He did an excellent job, resulting in a 10 times faster execution than the previous 2D parallel solver, which means that his new optimised code was 30 times faster than the original, in-house, serial solver.\par
Finally, in 2016, Maciej Kubat proposed a new version of the LBM solver.Firstly he converts the 2D parallel solver to a 3D parallel solver which entailed a major re-engineering of the code, from data containers to logic cycles \cite{maciej_thesis}. After first trying for a direct adaptation, he found that his code was too slow to produce meaningful solutions. After optimising his own code he was able to reach an almost one hundred times speed-up.
However, Kubat states that a lot can be done to improve his code, from boundary conditions to code readability and maintainability. 
\chapter{Methodology}
\section{In-house LBM solver}
The implementation of a 3D multiphase flow using the Lattice Botzmann method will be based on an existing in-house code. As such, the first step in developing the new solver is to acquire a good understanding of how the previous solver works, specifically, it’s structure and organisation, it’s input and output data, it’s performance and it’s depending software (if any). Furthermore, the previous code needs to be validated so as to provide and accurate foundation for the project being developed. Without this validation, it would prove to be an arduous work to discover whether the new solver is behaving correctly.
\subsection{Code organisation}
After spending some time analysing the LBM solver and debugging some of the core features, I was able to obtain a generous understanding of how the existing code is organised and where most of the necessary algorithms are located. The solver itself was very well documented and most of the solver’s features were written in a very user-friendly way. Because of this good work from the authors of this solver, the following flow chart, Figure \ref{fig:codeFlow}, was able to be made in a relatively small amount of time, which helps to gain a general understanding of the solver.

@@

@@

\subsection{Validation}
Now that the received code has been validated to perform under the same conditions as the final version used by the previous authors, we now have to verify whether the results that the received version is producing are the same as the final one. The validation of the results obtained throughout this project will be under the responsibility of Antonio González. As such, the details for this initial validation can be found in his thesis \cite{antonioThesis}.
\par This initial validation is essential for the development of a multiphase flow using the in-house solver. Without it, there would be no way of telling whether the solver would be functioning incorrectly due to a faulty algorithm implemented by us or if the solver was already producing wrong results. By doing so we can be sure that our code behaves correctly and that a proper result is achieved.

\chapter{Results and Discussion}
\chapter{Conclusions}

>>>>

\usepackage[utf8]{inputenc}
% Use the custom "cranfieldthesis" LaTeX style file. 
\usepackage{cranfieldthesis}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
% By default, LaTeX uses a serif font - these are traditionally thought to be
% easier to read.   If you'd prefer sans-serif, please uncomment the 
% following line.
%\renewcommand{\familydefault}{\sfdefault}

@@

\section{Lattice Boltzmann Method} \label{sec:lbmLit}
The Lattice Boltzmann method (LBM) is a mesoscopic scale approach to CFD and was first introduced as a Lattice-Gas Automata for the Navier-Stokes Equation \cite{gasautomata}. It is used to describe a fluid based on probabilities using the Maxwell-Boltzmann equation in the fluid’s equilibrium state \cite{lbm_springer}. 
\par In this method, we do not consider the individual characteristics of each particle. By grouping particles together in a D2Q9 (nodes containing 9 particles for 2D problems - Fig \ref{fig:d2q9})  or in a D3Q19 (nodes containing 19 particles for 3D problems - Fig \ref{fig:d3q19}) model, we can analyse the behaviour of the particles collectively  \cite{lbm_springer}.

@@

\subsection{Multiphase flow}
Multiphase flows represent the simultaneous flow of materials in different states (phases). As such, multiphase flows have an enormous spectrum of representation and stand for interactions between gas/solid flows, liquid/solid flows or even liquid/liquid flows with different chemical properties \cite{multiphase}.\par
These simulations present challenging problems because of inherent difficulties during modelling and the importance of engineering applications. However, the Lattice Boltzmann method provides an alternative for these simulations due to it's relative simplicity when compared with traditional Navier-Stokes equation \cite{chenDoolenMP}.
\par These flows have several applications in the scientific and industrial community, ranging from pourous media fluid interactions \cite{pourous} to flows containing gas bubbles dispersed in liquids \cite{gasBubble}. Furthermore, almost every processing technology must take multiphase flows into account, including cavitating pumps, papermaking and many others \cite{multiphase}.
\par Several implementation strategies exist for Multiphase flows, however, for the purpose of this thesis, we will first follow the simple ''two-color'' approach first proposed by Gunstensen et al \cite{gunstensen} \textcolor{red}{This is wrong, refer to RK. Also, explain the history and how to method was developed}.

@@

\section{Previous parallelisation works}
This thesis is a continuation of work that started some years ago in Cranfield University. As such, the work from the previous students needs to be analysed. \par
In 2014, Tamás Józsa and Máté Szőke, adapted two different in-house C and C++ codes into one single C code unifying the advantages of each one of the two original codes \cite{jozsa_thesis,szoke_thesis}. Józsa then parallelised the critical parts of the C code using CUDA and ran tests on the Fermi GPU Cluster from Cranfield, achieving a three times speed-up in general, with a peak of 15 times speed-up \cite{jozsa_thesis}. \par
Szőke proposes a CPU parallelisation approach using Unified Parallel C, which is a Partitioned Global Address Space language\cite{szoke_thesis}. This means that it is possible to use shared memory to compute the solution. However, the author verified that a local memory-based approach (like MPI) provided the best results. He also compared the results obtained with the ones obtained by Józsa on the CUDA approach. They found that to achieve the same speed-up as that of a single GPU card, one needs an entire workstation (16 threads in the case of Astral) \cite{szoke_thesis}.\par
In 2015, Ádám Koleszár continued the work and further optimised the parallel version of the LBM method using CUDA \cite{koleszar_thesis}. He did an excellent job, resulting in a 10 times faster execution than the previous 2D parallel solver, which means that his new optimised code was 30 times faster than the original, in-house, serial solver.\par
Finally, in 2016, Maciej Kubat proposed a new version of the LBM solver.Firstly he converts the 2D parallel solver to a 3D parallel solver which entailed a major re-engineering of the code, from data containers to logic cycles \cite{maciej_thesis}. After first trying for a direct adaptation, he found that his code was too slow to produce meaningful solutions. After optimising his own code he was able to reach an almost one hundred times speed-up.
However, Kubat states that a lot can be done to improve his code, from boundary conditions to code readability and maintainability. 
\chapter{Methodology}
\section{Lattice Boltzmann Method}
The Lattice Boltzmann method is a mesoscopic approach to solve several fluid dynamics problems, \textcolor{red}{as stated in Section}. It relies on a statistical description of the system via a distribution function $f$. This distribution function is responsible for predicting the number of molecules at a certain time, positioned between 2 points and with velocities between 2 values. This function is used in combination with a collision operator $\Omega$ and, with no external force being applied, represent the Boltzmann equation as
\begin{equation}\label{eq:boltzEq}
\frac{\partial f}{\partial t} + c \nabla f = \Omega
\end{equation}
giving us an advection equation. 
However, Eq. \ref{eq:boltzEq} is difficult to solve by itself. To offset this difficulty we need to approximate the collision operator with a simpler operator that will not introduce a significant error in the final solution. After solving this collision step, we then need to propagate the changes in each cell onto to their neighbours (streaming). We then solve boundary interactions within the nodes affected by the boundary spaces of the problem and finally we update the macroscopic values. These values will be fed into the collision step again where the process will repeat itself until the solution converges/diverges or the final number of iterations is reached.
\subsection{Discretization}
Before starting with the collision step, we must first discretize the domain into a mesh composed of various cells. Each of these cells will in turn affect and be affected by other cells, depending on the chosen speed model, as stated in Section \ref{sec:lbmLit}. As such, we will use a 2DQ9 arrangement for the 2D problems and a 3DQ19, which roughly translates into each cell interacting with another 8 or 18 cells (neighbours), depending on the problem's dimension.
\subsection{Collision}
There are a few collision models capable of approximating the collision operator $\Omega$, such as the BGKW \textcolor{red}{cite}, the TRT \textcolor{red}{cite} and the MRT  \textcolor{red}{cite}. However, for the purpose of this thesis, we will focus solely on the BGKW model as this will be the approximation used by the Multiphase model. \par
In the BGKW model, we can use the following equation to approximate $\Omega$
\begin{equation}\label{eq:bgkwEq}
\Omega = \omega \left( f^{eq} - f\right) = \frac{1}{\tau}\left( f^{eq} - f\right)
\end{equation}
where $\omega$ represents the collision frequency and $\tau$ stands for the relaxation factor. $f^{eq}$ represents the local equilibrium of the distribution function, which is known as the Maxwell-Boltzmann distribution function. Now, we can use Eq. \ref{eq:bgkwEq} in the discretized Boltzmann equation to obtain
\begin{equation}
f_i = \omega f^{eq}_i + (1 - \omega)f_i%\frac{\partial f_i}{\partial t} + c_i \nabla f_i = \frac{1}{\tau}\left( f^{eq}_i - f_i\right)
\end{equation}
which will be valid when following specific directions.
\subsection{Streaming}\label{sec:streaming}
After each node finishes calculating the changes brought on by the collision with other nodes, it is necessary to pass this information onto each neighbour. This streaming will take into account the directions imposed by the speed model. Figure \ref{fig:streaming} demonstrates how this step of the method works for the two dimension problem type.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\linewidth]{Resources/Images/streaming.png}
	\caption{Streaming step following a 2DQ9 model.}
	\label{fig:streaming}
\end{figure}

For 3D problems the streaming step follows the same strategy.
\subsection{Boundary update}\label{sec:boundary}
After each node updates their neighbour, the nodes located on the boundaries of the problem need to be updated depending on the boundary condition applied to them. Again, there are quite a few boundary conditions implemented in the in-house solver. However, the author will only go into detail in the boundary conditions used by the new Multiphase model.\par
\subsubsection{Wall}
	These boundary conditions are implemented with a bounce-back effect, where particles that are incoming into the solid wall bounce back into the domain. Applying the laws of conservation of mass and momentum, a particle incoming into the southern direction of the domain leads to  
	\begin{equation*}
		\begin{aligned}
			f_2 = f_4 \\
			f_5 = f_7 \\
			f_6 = f_8
		\end{aligned}
	\end{equation*}
	\subsubsection{Periodic boundary conditions} 
	With these boundaries, particles that exit the domain on one side will enter the domain on the opposite, creating an infinite corridor between opposite sides. As such, if a particle reaches the north boundary, their distribution function will be updated as follows
	\begin{equation*}
		\begin{aligned}
			f_4^N = f_4^S \\
			f_7^N = f_7^S \\
			f_8^N = f_8^S
		\end{aligned}
	\end{equation*}
	\subsubsection{Inlet}
	 Zou and He \textcolor{red}{cite} proposed a method for solving these boundary conditions. In this case, the boundary simulates a fluid flow entering the domain with a certain velocity. Applying the proposed equations to the BGKW model on a cell located in the north boundary yields:
	\begin{equation*}
		\begin{aligned}
			\rho &= \frac{1}{1+v} * \left(f_1 + f_3 + 2 * \left(f_2 + f_5 + f_6\right)\right) \\
			f_4 &= f_2 - \frac{2}{3}\rho v \\
			f_7 &= f_5 + \frac{1}{2}\left(f_1 - f_3\right) - \frac{1}{6}\left(\rho v\right) - \frac{1}{2}\left(\rho u\right) \\
			f_8 &= f_6 + \frac{1}{2}\left(f_1 - f_3\right) - \frac{1}{6}\left(\rho v\right) + \frac{1}{2}\left(\rho u\right)
		\end{aligned}
	\end{equation*}
\subsection{Macro-variables update}\label{sec:mpUpd}
Finally, after updating the distribution function of every cell in the domain, we can update the macroscopic variables $\rho$, $u$ and $v$.
From Eq. \ref{eq:boltzEq}, we can define the macroscopic values for density and velocity as
\begin{equation}\label{eq:densityUp}
\rho (r,t) = \int  mf(r,c,t)dc
\end{equation}
\begin{equation}\label{eq:velocityUp}
\rho (r,t) u(r,t) = \int  mcf(r,c,t)dc
\end{equation}
where $m$ represents the molecular mass.
\section{Color Gradient Model}
Models implementing multiphase flows for the Lattice Boltzmann method can generally be classified into several different categories. For the purpose of this thesis, we will focus on the Rothman-Keller methods, specifically on the model proposed by Reis and Phillips \textcolor{red}{cite}, modified using Latva-Kokko's recoloring operator, as proposed by Leclaire et al. \textcolor{red}{cite leclaire}. \par
In this model, two immiscible fluids are simulated under the LBM, a red fluid and a blue fluid, that only interact with each other in the interface between them. These fluids are associated to their own distribution functions, which implies that the total memory used by the in-house solver is likely to increase. Furthermore, the collision step will need to be created from scratch since this step introduces two new operators, while the standard collision step will also suffer some changes to account for the two fluids. \par
The fluid's distribution function now becomes
\begin{equation}\label{eq:cgDistrib}
f_{i}^{k}\left(x+c_i, t+1\right) = f_{i}^{k}\left(x,t\right) + \Omega_i^{k}
\end{equation}
where $k$ stands for the fluid, either red or blue.$\Omega_i^{k}$ is the result of the combination of the 3 sub-steps present in the new collision step and is given by
\begin{equation}\label{eq:cgColl}
\Omega_i^{k} = \left(\Omega_i^{k}\right)_{\left(3\right)} \left(\left(\Omega_i^{k}\right)_{\left(1\right)} + \left(\Omega_i^{k}\right)_{\left(2\right)}\right)
\end{equation}
\subsection{Single-phase collision operator}
The first sub-step is similar to the collision operator used in the standard BGKW model, introduced in Eq. \ref{eq:bgkwEq}. However, a new operator, $\phi$ is introduced in the calculation of the local equilibrium distribution function. 
\begin{equation}\label{eq:newFEQ}
f_i^{k\left(eq\right)} = \rho_k\left(\phi_i^k + W_i\left(3c_i \cdot u + \frac{9}{2}\left(c_i \cdot u\right)^2 - \frac{3}{2}u^2 \right)\right)
\end{equation}
where $\phi_i^k$ is given by 
\begin{itemize}
	\item $\alpha_k$ for $i = 0$
	\item $\frac{\left(1-\alpha_k\right)}{5}$ for $i = {1,2,3,4}$
	\item $\frac{\left(1-\alpha_k\right)}{20}$ for $i = {5,6,7,8}$
\end{itemize}
\subsection{Perturbation operator}
After computing the single-phase collision operator we need to add the result of the perturbation operator before passing it on to the recoloring sub-step. This is where we simulate the surface tension between the fluids and ensure that pressure difference is in equilibrium.\par
First, we must calculate the color gradient term, which is defined by
\begin{equation}\label{eq:colorGrad}
F = \sum_i c_i \left(\rho_r\left(x+c_i\right) - \rho_B\left(x+c_i\right)\right)
\end{equation}
which is 4th order accurate. Then, we include $F$ in the calculation of the perturbation term, defined by
\begin{equation}\label{eq:pert}
\left(\Omega_i^{k}\right)_{\left(2\right)}\left(F_i^k\right) = F_i^k + \frac{A_k}{2}\norm{F}\left(W_i\frac{\left(F\cdot c_i\right)^2}{\norm{F}^2} - B_i\right)
\end{equation}
\subsection{Recoloring operator}
The last step in the collision step is calculating the recoloring operator. This operator guarantees that the fluids remain immiscible and also controls the amount of a fluid sent to it's corresponding region. The operator can be defined as
\begin{equation}
\left(\Omega_i^{k}\right)_{\left(3\right)}\left(F_i^k\right) = \frac{\rho_k}{\rho}F_i + \beta\frac{\rho_r \rho_b}{\rho^2}\cos \left(\phi_i\right)\sum_k F_i^{k\left(eq\right)}\left(\rho_k,0,\alpha_k\right)
\end{equation}
where $\beta$ is a free parameter between 0 and 1 that influences the thickness of the interface and $\cos \left(\phi_i\right)$ is the cosine of the angle between the color gradient $F$ and the direction $c$.
\subsection{Streaming and Boundary conditions}
The streaming and boundary conditions in this model are analogous to the ones used in the LBM method, as defined in Sections \ref{sec:streaming} and \ref{sec:boundary}. To compute them, we simply need to repeat these steps for the distribution function of each fluid.
\subsection{Macro-variables update}
This step is also similar to the one defined in Section \ref{sec:mpUpd}. The main difference is that $\rho$ is the result of the sum of the densities of each fluid and $F_i$ is also the sum of the distribution functions of each fluid. 
\section{Meshes}
For the LBM in-house solver we need to prepare structured meshes capable of representing the domain. These meshes allow us to discretize the functions and solve them according to the representation of space. Each mesh is composed of equidistant small spaces that represent the cells used in the equations. To this end, software capable of generating meshes is needed, that can both create the mesh according to the number of nodes required and specify the type of boundary conditions to be used. G. Abbruzzese developed software capable of generating meshes with these requirements \textcolor{red}{cite} which has been used in the previous thesis over the LBM in-house solver. \par
However, only the 3D lattice generator still exists, which means that we are restricted to the 2D meshes that already exist. Luckily, these meshes are sufficient to simulate the requirements of all of the proposed test cases. The boundaries however will need to be hard-coded to match the initial requirements and to adequately simulate the 2D test cases.
\subsection{Used meshes}
The test cases for the Color Gradient model are based on a simple square for 2D or a cube for 3D problems. However, the boundaries are specific to each case. All of them use periodic boundaries on the East and West directions and most of them also use them for the North and South directions, the exception being for the Couette flow where the North boundary is an Inlet and the South boundary acts a solid wall.
\section{CUDA programming}
As stated before, CUDA is a parallel programming platform introduced by NVIDIA. One of it's most appealing features is the fact that it is integrated into the well known programming languages C/C++, making it easy for programmers to start developing parallel code in a familiar environment. However, there are some details that one must know before being able to take benefit of the computational capabilities of GPUs. Note that the host is referring to the CPU and the device is referring to the GPU.
\subsection{Thread arrangement}
As is the case with several other parallel programming platforms, CUDA's base parallel agents are threads. CUDA allows for a massive number of threads running concurrently. However, accessing each thread is not as straightforward as one would hope. Figure \ref{fig:threads} show how threads are organized inside CUDA's memory model.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\linewidth]{Resources/Images/threadmapping.png}
	\caption{Thread arrangement in CUDA platform.}
	\label{fig:threads}
\end{figure}

Threads are organized into blocks and blocks are organized in the grid. Both the grid of blocks and the blocks themselves can be of 1, 2 or 3 dimensions. CUDA provides a very useful way of obtaining the index of each thread. We can easily obtain the thread's index inside the block as well as the block's index in the grid. However, some math is needed to obtain the thread's global index, which is very useful for memory management, such as access to an array's element. Fortunately, this is an easy task since we can access all the needed information with threadIdx, blockDim, blockIdx and gridDim.

\subsection{Memory structures}
Since CUDA code runs on GPUs, it is worth showing what the internal architecture of a NVIDIA GPU looks like. 
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\linewidth]{Resources/Images/cudaMemory.png}
	\caption{Memory arrangement in CUDA compliant GPU.}
	\label{fig:memCuda}
\end{figure}

Figure \ref{fig:memCuda} shows 4 different types of memory available to each thread
\begin{itemize}
	\item Registers
	\item Shared memory
	\item Global memory
	\item Constant memory
\end{itemize}

There is also one more memory type, texture memory. However, this memory type will not be used in the scope of this thesis, so the author will not go into detail about them.
\subsubsection{Registers}
Registers are the fastest type of memory available to a thread. Each thread's register is only visible to itself and so every variable using this type of memory is visible only to the thread. This means that it is not possible to use data directly from the host on these structures. Also, this data is non-transferable between threads. \par Registers work just like local memory for each thread, the difference is that local memory is comparable to global memory in terms of speed since it is not allocated directly on the GPU chip but is instead an abstraction of global memory. Local memory will only be used if the compiler determines that the thread's register size is not large enough to hold the thread's local memory.
\subsubsection{Shared memory} 
Shared memory is the second fastest type of memory available to a thread. As opposed to registers, shared memory is, as the name suggests, shared between threads residing in the same block. This means that we can have threads cooperating with each other when using this type of memory. However, memory access needs to be properly managed, otherwise degrading memory access speed when bank conflicts occur. Bank conflicts means that two or more threads are trying to access the same memory address, and when this happens the memory access is serialized between the conflicting threads.\par
The nature of this memory makes it an excellent candidate for cases where threads require information from other threads to continue with their work, for example in a matrix vector multiplication. In these cases, a proper management of shared memory within the block can greatly boost the algorithms performance. As such, this a memory structure that should be used whenever threads require cooperation since with no bank conflicts shared memory can reach the speed of register memory.
\subsubsection{Constant memory}
This a read-only type of memory that is stored in the cache, making access to it faster than global memory. This is the first memory type that can transfer data between the host and the device since before using in the device data needs to be properly initialized within the host code. Cached memory has the benefit that a single read can be broadcast to other 15 threads, and also that consecutive reads from the same address space will not incur any additional memory traffic \textcolor{red}{cite cuda}. However, we can only benefit from a performance gain when using constant memory if we pay attention to the warps inside a block, since the memory read is broadcast to a half-warp (more on this later).
\subsubsection{Global memory}
Finally, we reach the slowest yet most versatile memory in a CUDA compliant GPU. Global memory has, as the name suggests, a global scope, meaning that every thread inside the grid can access it, as is the case with constant memory. However, it is not limited to read-only access as all threads can also write data onto it. This is the only other data that can be transferred between host and device (apart from texture memory) but it has the disadvantage of being the slowest memory type. To boost performance, data in global memory will usually be read into a faster memory type like registers or shared memory before having computations performed over it.

\subsection{Warps}
A warp is a group of 32 threads inside a block. Each block, when it is created, is assigned to a Streaming Multiprocessor that contain 32 processing cores \textcolor{red}{cite webinar}. The threads inside a warp are truly concurrent, meaning that at a given time each thread will execute the same instruction concurrently (lock-step fashion). Therefore, warp-awareness is essential to write highly-optimised code \textcolor{red}{cite salvatore}. One must guarantee that each thread inside the warp will follow the same control path, otherwise risk wasting the full potential of the warps concurrency by achieving warp divergence.

\chapter{Results and Discussion}
\section{In-house LBM solver}
The implementation of a 3D multiphase flow using the Lattice Botzmann method will be based on an existing in-house code. As such, the first step in developing the new solver is to acquire a good understanding of how the previous solver works, specifically, it’s structure and organisation, it’s input and output data, it’s performance and it’s depending software (if any). Furthermore, the previous code needs to be validated so as to provide and accurate foundation for the project being developed. Without this validation, it would prove to be an arduous work to discover whether the new solver is behaving correctly.
\subsection{Code organisation}
After spending some time analysing the LBM solver and debugging some of the core features, I was able to obtain a generous understanding of how the existing code is organised and where most of the necessary algorithms are located. The solver itself was very well documented and most of the solver’s features were written in a very user-friendly way. Because of this good work from the authors of this solver, the following flow chart, Figure \ref{fig:codeFlow}, was able to be made in a relatively small amount of time, which helps to gain a general understanding of the solver.

@@

\subsection{Data representation}
The original code mostly handles multidimensional arrays, following the same logic from the 2D solver onto the 3D version. As such, the arrays either store the macroscopic value of each node (h * m * n in 3D) or the coefficient for the microscopic values of each lattice (19 * h * m * n in 3D) . However, when using CUDA kernels, we will often have the need to copy data from the host (CPU) to the device (GPU), which can prove to be a tiring and inefficient task when dealing with 2, 3 or even 4 dimension arrays. To solve this problem, the previous authors flattened every array, which means that all arrays are represented as one dimensional arrays. Because of this, some calculations are required to access the desired index, \textcolor{red}{as demonstrated by Fig.}
\par These arrays are stored in a row-major fashion and the lattices are grouped together in 9 or 19 arrays, depending on the dimension of the problem, meaning that neighbours are separated from each other by m * n or h * m * n elements.

@@

\subsection{Validation}
Now that the received code has been validated to perform under the same conditions as the final version used by the previous authors, we now have to verify whether the results that the received version is producing are the same as the final one. The validation of the results obtained throughout this project will be under the responsibility of Antonio González. As such, the details for this initial validation can be found in his thesis \cite{antonioThesis}.
\par This initial validation is essential for the development of a multiphase flow using the in-house solver. Without it, there would be no way of telling whether the solver would be functioning incorrectly due to a faulty algorithm implemented by us or if the solver was already producing wrong results. By doing so we can be sure that our code behaves correctly and that a proper result is achieved.
\chapter{Conclusions}
